Self-interacting dark matter model without dark energy in cosmology

### 2024-12-08 article/article.tex
qso constraints

\documentclass{aa}
\usepackage{graphicx}
\usepackage{txfonts}

@@

   If $n=0$, the global density evolution can be transformed as
   \begin{equation}
      \rho(a)=C_1+a^{-3}\frac{e^{-2\kappa C_1(t-t_0)}}{C_2-\kappa Z(t)},
   \end{equation}
   where $Z(t)=\int_{t_0}^ta^{-3}e^{-2\kappa C_1(t'-t_0)}\mathrm{d}t'$(\cite{refId0}).
   Using the usual unit-less $\Omega$-like parameterization, we obtain that
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{1,0}+\frac{\Omega_{1,0}\Omega_{2,0}a^{-3}e^{-2\kappa C_1(t-t_0)}}
      {\Omega_{1,0}+\kappa C_1\Omega_{2,0}Z(t)},
   \end{equation}
   where $\Omega_{1,0}=8\pi GC_1/3H_0^2$ and $\Omega_{2,0}=8\pi G/3H_0^2C_2$,
   which related to $\Omega_{\Lambda}$ and $\Omega_m$ in the $\Lambda$CDM model,
   respectively.
   From Eq.({\ref{eq:2}}), we can also give the mass of the DM particle
   related to the range of $\kappa C_1$(in the unit of Gyr${}^{-1}$)
   \begin{equation}
      M_x=\frac{3.325\times10^{-12}}{\kappa C_1}
      \frac{\langle\sigma u\rangle}{10^{-23}}h^2(1-\Omega_{2,0})\,\text{GeV},\label{eq:9}
   \end{equation}
   where $h\equiv H_0/[100\text{km/s/Mpc}]$.

@@

   We use the fiducial SN Ia magnitude ($M_b$) determined from SH0ES 2021 Cepheid 
   host distances(\cite{Riess_2022}), which gives the $\mu_\text{data}$ and constrains $H_0$
   in advance. To eliminate the influence of $M_b$, we give the likelihood function
   as \begin{equation}
      \chi_{\text{SNe}}^2=A-\frac{B^2}{C}+\ln\left(\frac{C}{2\pi}\right),
   \end{equation}
   where $A=\sum_{i=1}^{1701}(\mu_{\text{th}}-\mu_{\text{data}})^2/\sigma_i^2,
   B=\sum_{i=1}^{1701}(\mu_{\text{th}}-\mu_{\text{data}})/\sigma_i^2,C=\sum_i^{1701}1/\sigma_i^2$,
   the distance modulus is $\mu=5\log_{10}(d_L/\text{Mpc})+25$, and the
   luminosity distance $d_L$ can be given as a function of redshift $z$
   \begin{equation}
      d_L=(1+z)\int_0^z\frac{c\mathrm{d}z'}{H(z')}.
   \end{equation}
   However, the Eq.(\ref{eq:9}) just need the $H_0$ to caculate the $M_x$
   and we would still use the simple likelihood function as
   \begin{equation}
      \tilde{\chi}_{\text{SNe}}^2=A=\sum_{i=1}^{1701}\frac{(\mu_{\text{th}}-\mu_{\text{data}})^2}{\sigma_i^2}.
   \end{equation}
   Apparently these two functions give the same constraints in $\Omega_{2,0}$
   and $\kappa C_1$.

@@

   The Baryon acoustic oscillation method (BAO) provides a key cosmological probe
   sensitive to the cosmic expansion history with well-controlled systematics.
   We use two BAO data sets from the SDSS(\cite{PhysRevD.103.083533}) and DESI 2024(\cite{desicollaboration2024desi2024vicosmological}),
   which are given at Table \ref{tab:2} and Table \ref{tab:3}, respectively.

>>>>

\documentclass{aa}
\usepackage{graphicx}
\usepackage{txfonts}
\graphicspath{{./pictures/}}

@@

   If $n=0$, the global density evolution can be transformed as
   \begin{equation}
      \rho(a)=C_1+a^{-3}\frac{e^{-2\kappa C_1(t-t_0)}}{C_2-\kappa Z(t)},
   \end{equation}
   where $Z(t)=\int_{t_0}^ta^{-3}e^{-2\kappa C_1(t'-t_0)}\mathrm{d}t'$(\cite{refId0}).
   Using the usual unit-less $\Omega$-like parameterization, we obtain that
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{1,0}+\frac{\Omega_{1,0}\Omega_{2,0}a^{-3}e^{-2\kappa C_1(t-t_0)}}
      {\Omega_{1,0}+\kappa C_1\Omega_{2,0}Z(t)},\label{eq:8}
   \end{equation}
   where $\Omega_{1,0}=8\pi GC_1/3H_0^2$ and $\Omega_{2,0}=8\pi G/3H_0^2C_2$,
   which related to $\Omega_{\Lambda}$ and $\Omega_m$ in the $\Lambda$CDM model,
   respectively.
   From Eq.({\ref{eq:2}}), we can also give the mass of the DM particle
   related to the range of $\kappa C_1$(in the unit of Gyr${}^{-1}$)
   \begin{equation}
      M_x=\frac{3.325\times10^{-12}}{\kappa C_1}
      \frac{\langle\sigma u\rangle}{10^{-23}}h^2(1-\Omega_{2,0})\,\text{GeV},\label{eq:9}
   \end{equation}
   where $h\equiv H_0/[100\text{km/s/Mpc}]$.

@@

   We use the fiducial SN Ia magnitude ($M_b$) determined from SH0ES 2021 Cepheid 
   host distances(\cite{Riess_2022}), which gives the $\mu_\text{data}$ and constrains $H_0$
   in advance. To eliminate the influence of $M_b$, we give the $\chi^2$
   as \begin{equation}
      \chi_{\text{SNe}}^2=A-\frac{B^2}{C}+\ln\left(\frac{C}{2\pi}\right),
   \end{equation}
   where $A=\sum_{i=1}^{1701}(\mu_{\text{th}}-\mu_{\text{data}})^2/\sigma_i^2,
   B=\sum_{i=1}^{1701}(\mu_{\text{th}}-\mu_{\text{data}})/\sigma_i^2,C=\sum_i^{1701}1/\sigma_i^2$,
   the distance modulus is $\mu=5\log_{10}(d_L/\text{Mpc})+25$, and the
   luminosity distance $d_L$ can be given as a function of redshift $z$
   \begin{equation}
      d_L=(1+z)\int_0^z\frac{c\mathrm{d}z'}{H(z')}.
   \end{equation}
   However, the Eq.(\ref{eq:9}) just need the $H_0$ to caculate the $M_x$
   and we would still use the simple likelihood function as
   \begin{equation}
      \tilde{\chi}_{\text{SNe}}^2=\Delta^{\text{T}}C^{-1}\Delta,
   \end{equation}
   where $\Delta=(\mu_{\text{th}}-\mu_{\text{data}})$ and $C^{-1}$ is the inverse of the
   covariance matrix of the SNe Ia data.

   $B=\Delta^{\text{T}}C^{-1}$ and $C=\text{sum}(C^{-1})$.

\subsection{Quasar}

   The quasar gives a higher redshift

@@

   The Baryon acoustic oscillation method (BAO) provides a key cosmological probe
   sensitive to the cosmic expansion history with well-controlled systematics.
   We use two BAO data sets from the SDSS(\cite{PhysRevD.103.083533}) and DESI 2024(\cite{desicollaboration2024desi2024vicosmological}),
   which are given at Table \ref{tab:2} and Table \ref{tab:3}, respectively.
   The redshift is up to 2.33 both in the SDSS and the DESI 2024 dataset.

   The $\chi^2$ function for the BAO data is defined as
   \begin{equation}
      \chi_{\text{BAO}}^2=\sum_i\frac{(D_{\text{th}}/r_{\text{d}}-D_{\text{data}}/r_{\text{d}})^2}{\sigma_i^2},
   \end{equation}
   where $D$ refers to $D_{\text{M}}$, $D_{\text{H}}$, or $D_{\text{V}}$, which ara given as
   \begin{eqnarray}
      D_{\text{M}}(z)&=&c\int_0^z\frac{\text{d}z'}{H(z')},\\
      D_{\text{H}}(z)&=&\frac{c}{H(z)},\\
      D_{\text{V}}(z)&=&\left[zD_{\text{M}}^2(z)D_{\text{H}}(z)\right]^{1/3},\\
   \end{eqnarray}
   and $r_{\text{d}}$ is the sound horizon at the drag epoch, which is given as
   \begin{equation}
      r_{\text{d}}=\int_{z_{\text{drag}}}^{\infty}\frac{c_s\text{d}z'}{H(z')}.
   \end{equation}

   However, the Eq.(\ref{eq:8}) just have a stiff point when $z\to\infty$, so the IDM
   model can not give a constraint to the $r_{\text{d}}$ and we just try to use the cross
   parameter $r_{\text{d}}h$ to give the constraints.
